Multiple quantum NMR in solids as a method of a determination of Wigner--Yanase skew information

### 2021-02-17 manuscript.tex
Add refferece and fix bibitem typos

MQ NMR methods are widely used for solving problems of quantum informatics~\cite{16,17}. The MQ NMR experiment consists of four distinct periods of time Fig.~(\ref{fig:1}): preparation $(\tau)$, evolution $(t_1)$, mixing $(\tau)$, and detection $(t_2)$~\cite{9}.
MQ NMR coherences are created by a periodic multipulse sequence, consisting of
$\pm$-pulses, irradiating the system of the preparation period~\cite{9}.
If the inverse period of the multipulse sequence significantly exceeds the local dipolar field (in frequency units)~\cite{18}
MQ NMR dynamics can be described by the averaged nonsecular two--spin/two--quantum Hamiltonian~$H_{MQ}$
%
\begin{equation} \label{eq:1}
        H_{MQ} = H^{(+2)} + H^{(-2)} , \quad
        H^{(\pm 2)} = -\frac{1}{2} \sum_{j<k} D_{jk} I_{j} ^\pm I_k^\pm,
\end{equation}
%
where $D_{jk}$ is the coupling constant between spins $j$ and $k$, and $I_{j} ^+ I_k ^-$ are the raising and lowering operators of spin $j$.
On the mixing period the spin system is irradiated by the multiple sequence with $\pm y$--pulses.
As a result, the averaged nonsecular two-spin/two quantum Hamiltonian on the mixing period equals $(-H_{MQ})$~\cite{9}.

@@

It is convenient to expand the density matrix $\rho_\mathrm{pre}(\tau)$ in series as~\cite{19}
\begin{equation}
    \label{eq:6}
        \rho_\mathrm{pre}(\tau,\beta) = \sum_n \rho_{\mathrm{pre},n}(\tau,\beta),
\end{equation}
where $\rho_\mathrm{pre}(\tau,\beta)$ is the contribution to the density matrix $\rho_\mathrm{pre}(\tau,\beta)$ from the MQ coherence of the n--th order.
Then the resulting signal $G(\tau,\phi)$ of the MQ NMR~\cite{9} can be rewritten as
\begin{equation} \label{eq:7}
    G(\tau, \phi) = \sum_n e^{in\phi}
        Tr\left\{\rho_{\mathrm{pre},n}(\tau,\beta)
        \rho_{\mathrm{pre},-n}(\tau,\beta) \right\},
\end{equation}
where we took into account that
\begin{equation}
    \label{eq:8}
        [I_z, \rho_{\mathrm{pre},n}] = n \rho_{\mathrm{pre},n}
\end{equation}
The normalized intensities of the MQ NMR coherences can be determined as follows
\begin{equation}
    \label{eq:9}
        J_n(\tau,\beta)= \frac{Tr\left\{\rho_{\mathrm{pre},n}(\tau,\beta)
            \rho_{\mathrm{pre},-n}(\tau,\beta)\right\}}
                {Tr(\rho^2_\mathrm{eq})}
\end{equation}
As was shown in~\cite{7},
\begin{equation}
    \label{eq:10}
        Tr(\rho_\mathrm{eq}^2) = \frac{2^N ch^N (\beta)}{Z^2}
\end{equation}
It was also shown that
\begin{equation}
    \label{eq:11}
        \sum_n J_n(\tau,\beta) = 1
\end{equation}
The second moment (dispersion) $M_2(\tau,\beta)$ of the distribution of the MQ NMR coherences $J_n (\tau,\beta)$ can be calculated from Eq.~(\ref{eq:7}) according to~\cite{20}
\begin{equation}
    \label{eq:12}
        M_2(\tau,\beta) = -\frac{1}{G_2(\tau,0)}
            \frac{d^2 G(t,\beta)}{dt^2}\bigg|
        _{t=0}
\end{equation}
Using Eqs.~(\ref{eq:7},\ref{eq:8},\ref{eq:12}) one can obtain
\begin{equation}
    \label{eq:13}
        M_2 (\tau,\beta) = \sum_n n^2 J_n(\tau,\beta)
\end{equation}
The low bound on the quantum Fisher information coincides with the double second moment of Eq.~(\ref{eq:13})~\cite{6,8}.
As a result, an analysis of the temperature dependence of the second moment $M_2(\tau,\beta)$ of the distribution of the intensities of the MQ NMR coherences allows us to obtain the number of entangled spins at different temperatures~\cite{7}.
In the following~Sec.~\ref{sec:3} we demonstrate that the Wigner--Yananse skew information is also connected with the second momentum $M_2(\tau,\beta)$
and can be useful for an investigation of the many-spin entanglement.

@@

The Wigner--Yanase skew information $I_{WY}(\rho(\tau,\beta),I_z)$ and the Fisher information $I_F(\rho(\tau,\beta),I_z)$ can be used for an investigation of the many-spin entanglement.
The informations are connected by the following restriction~\cite{3}
%
\begin{equation} \label{eq:22}
    I_{WY}\left(\rho(\tau,\beta), I_z\right)
    \leq I_F\left(\rho(\tau,\beta), I_z\right)
    \leq 2I_{WY}\left(\rho(\tau,\beta), I_z\right).
\end{equation}
%
For the comparison we used the model~\cite{21} of a nonspherical nanopore filled with a gas of spin-carrying atoms (for example, xenon) or molecules in a strong external magnetic field.
The restrictions (22) allow us to hope that the obtained results of the number of entangled spins are not strongly different.
This model allows to investigate the many-spin entanglement in the spin system consisting of hundreds of nuclear spins~\cite{7}.

@@

The analogous investigation was conducted on the model of the proton zigzag chain in a single crystal of hambergite~\cite{22}.
Here the close results on the many--spin entanglement were obtained for both used informations for the system consisting of $6\div 12$ spins.

@@

\bibitem{13} T.Macri, A. Smerzi, L.Pezze, Phys.Rev F94, 010102 (2016)
\bibitem{14} M.Gartner, J.G.Bohnet, A.Safavi-Naini, M.L. Wall, J.J. Bollinger, A.M.Rey, Nat.Phys. 13,781 (2017)
\bibitem{15} G.A.Bochkin, S.G.Vesil'ev, S.I. Doronin, E.I.Kuznetsova,I.D.Lazarev,E.B.Fel'dman, Appl.Magn.Reson. 51,667-678 (2020)
\bibitem{16} E.B.Fel'dman, A.N. Pykov, A.I.Zenchuk, Philos. Trans. R. Soc. London A370, 4690 (2012)
\bibitem{17} G.B.Furman, V.M.Meerovich, V.L.Sokoovsky, Phys. Rev. A.78 042301 (2008)
\bibitem{18} M.Goldman. Spin temperature and nuclear magnetic resonance in solids, Oxford,UK, Clarendon Press. 1970.
\bibitem{19} E.B.Fel'dman, S.Lacelle,Chem.Phys.Lett. 253, 27 (1996)
\bibitem{20} A.Abragam, The Princioles of Nuclear Magnetism, Clarendon. Oxford. 1961
\bibitem{21} J. Baugh, A. Kleinhammes, D. Han, Q. Wang, Y.Wn, Science 294, 1505 (2001)
\bibitem{22} G.A. Bochkin, E.B. Fel'dman, E.B. Kuznetsova, I.D. Lasarev, S.G. Vasil'ev, V.I.Volkov, J. Magn.Reson. 319, 106816 (2020)
\end{thebibliography}

>>>>

MQ NMR methods are widely used for solving problems of quantum informatics~\cite{16,17}. The MQ NMR experiment consists of four distinct periods of time Fig.~(\ref{fig:1}): preparation $(\tau)$, evolution $(t_1)$, mixing $(\tau)$, and detection $(t_2)$~\cite{9}.
MQ NMR coherences are created by a periodic multipulse sequence, consisting of
$\pm$-pulses, irradiating the system of the preparation period~\cite{9}.
If the inverse period of the multipulse sequence significantly exceeds the local dipolar field (in frequency units)~\cite{18}
MQ NMR dynamics can be described by the averaged nonsecular two--spin/two--quantum Hamiltonian~$H_{MQ}$~\cite{19}
%
\begin{equation} \label{eq:1}
        H_{MQ} = H^{(+2)} + H^{(-2)} , \quad
        H^{(\pm 2)} = -\frac{1}{2} \sum_{j<k} D_{jk} I_{j} ^\pm I_k^\pm,
\end{equation}
%
where $D_{jk}$ is the coupling constant between spins $j$ and $k$, and $I_{j} ^+ I_k ^-$ are the raising and lowering operators of spin $j$.
On the mixing period the spin system is irradiated by the multiple sequence with $\pm y$--pulses.
As a result, the averaged nonsecular two-spin/two quantum Hamiltonian on the mixing period equals $(-H_{MQ})$~\cite{9}.

@@

It is convenient to expand the density matrix $\rho_\mathrm{pre}(\tau)$ in series as~\cite{20}
\begin{equation}
    \label{eq:6}
        \rho_\mathrm{pre}(\tau,\beta) = \sum_n \rho_{\mathrm{pre},n}(\tau,\beta),
\end{equation}
where $\rho_\mathrm{pre}(\tau,\beta)$ is the contribution to the density matrix $\rho_\mathrm{pre}(\tau,\beta)$ from the MQ coherence of the n--th order.
Then the resulting signal $G(\tau,\phi)$ of the MQ NMR~\cite{9} can be rewritten as
\begin{equation} \label{eq:7}
    G(\tau, \phi) = \sum_n e^{in\phi}
        Tr\left\{\rho_{\mathrm{pre},n}(\tau,\beta)
        \rho_{\mathrm{pre},-n}(\tau,\beta) \right\},
\end{equation}
where we took into account that
\begin{equation}
    \label{eq:8}
        [I_z, \rho_{\mathrm{pre},n}] = n \rho_{\mathrm{pre},n}
\end{equation}
The normalized intensities of the MQ NMR coherences can be determined as follows
\begin{equation}
    \label{eq:9}
        J_n(\tau,\beta)= \frac{Tr\left\{\rho_{\mathrm{pre},n}(\tau,\beta)
            \rho_{\mathrm{pre},-n}(\tau,\beta)\right\}}
                {Tr(\rho^2_\mathrm{eq})}
\end{equation}
As was shown in~\cite{7},
\begin{equation}
    \label{eq:10}
        Tr(\rho_\mathrm{eq}^2) = \frac{2^N ch^N (\beta)}{Z^2}
\end{equation}
It was also shown that
\begin{equation}
    \label{eq:11}
        \sum_n J_n(\tau,\beta) = 1
\end{equation}
The second moment (dispersion) $M_2(\tau,\beta)$ of the distribution of the MQ NMR coherences $J_n (\tau,\beta)$ can be calculated from Eq.~(\ref{eq:7}) according to~\cite{21}
\begin{equation}
    \label{eq:12}
        M_2(\tau,\beta) = -\frac{1}{G_2(\tau,0)}
            \frac{d^2 G(t,\beta)}{dt^2}\bigg|
        _{t=0}
\end{equation}
Using Eqs.~(\ref{eq:7},\ref{eq:8},\ref{eq:12}) one can obtain
\begin{equation}
    \label{eq:13}
        M_2 (\tau,\beta) = \sum_n n^2 J_n(\tau,\beta)
\end{equation}
The low bound on the quantum Fisher information coincides with the double second moment of Eq.~(\ref{eq:13})~\cite{6,8}.
As a result, an analysis of the temperature dependence of the second moment $M_2(\tau,\beta)$ of the distribution of the intensities of the MQ NMR coherences allows us to obtain the number of entangled spins at different temperatures~\cite{7}.
In the following~Sec.~\ref{sec:3} we demonstrate that the Wigner--Yananse skew information is also connected with the second momentum $M_2(\tau,\beta)$
and can be useful for an investigation of the many-spin entanglement.

@@

The Wigner--Yanase skew information $I_{WY}(\rho(\tau,\beta),I_z)$ and the Fisher information $I_F(\rho(\tau,\beta),I_z)$ can be used for an investigation of the many-spin entanglement.
The informations are connected by the following restriction~\cite{3}
%
\begin{equation} \label{eq:22}
    I_{WY}\left(\rho(\tau,\beta), I_z\right)
    \leq I_F\left(\rho(\tau,\beta), I_z\right)
    \leq 2I_{WY}\left(\rho(\tau,\beta), I_z\right).
\end{equation}
%
The restrictions (22) allow us to hope that the obtained results of the number of entangled spins are not strongly different.
For the comparison we used the model~\cite{22} of a nonspherical nanopore filled with a gas of spin-carrying atoms (for example, xenon) or molecules in a strong external magnetic field.
This model allows to investigate the many-spin entanglement in the spin system consisting of hundreds of nuclear spins~\cite{7}.

@@

The analogous investigation was conducted on the model of the proton zigzag chain in a single crystal of hambergite~\cite{23}.
Here the close results on the many--spin entanglement were obtained for both used informations for the system consisting of $6\div 12$ spins~\cite{15}.

@@

\bibitem{13} T.Macri, A. Smerzi, L.Pezze, Phys.Rev A94, 010102 (2016)
\bibitem{14} M.Gartner, J.G.Bohnet, A.Safavi-Naini, M.L. Wall, J.J. Bollinger, A.M.Rey, Nat.Phys. 13,781 (2017)
\bibitem{15} G.A.Bochkin, S.G.Vesil'ev, S.I. Doronin, E.I.Kuznetsova,I.D.Lazarev,E.B.Fel'dman, Appl.Magn.Reson. 51,667-678 (2020)
\bibitem{16} E.B.Fel'dman, A.N. Pykov, A.I.Zenchuk, Philos. Trans. R. Soc. London A370, 4690 (2012)
\bibitem{17} G.B.Furman, V.M.Meerovich, V.L.Sokolovsky, Phys. Rev. A.78 042301 (2008)
\bibitem{18} M.Goldman. Spin temperature and nuclear magnetic resonance in solids, Oxford,UK, Clarendon Press. 1970.
\bibitem{19} S.I.Doronin, I.I.Maksimov, E.B.Fel’dman, J. Exp. Theor. Phys.91, 597 (2000)
\bibitem{20} E.B.Fel'dman, S.Lacelle,Chem.Phys.Lett. 253, 27 (1996)
\bibitem{21} A.Abragam, The Principles of Nuclear Magnetism, Clarendon. Oxford. 1961
\bibitem{22} J. Baugh, A. Kleinhammes, D. Han, Q. Wang, Y.Wn, Science 294, 1505 (2001)
\bibitem{23} G.A. Bochkin, E.B. Fel'dman, E.B. Kuznetsova, I.D. Lasarev, S.G. Vasil'ev, V.I.Volkov, J. Magn.Reson. 319, 106816 (2020)
\end{thebibliography}
